---
title: "Data607 Assignment"
author: "David Apolinar, Anthony Muñoz, Christina Valore"
date: "2/14/2019"
output: 
  html_document:
   toc: true
   toc_float: true
---
# Overview
Team: Anthony, Christina, David

Our team decided to use the following libraries to clean and select the data:

  stringr
  magrtitr
  dplyer
  DT
# Data Cleansing

In this section, we used the built-in R csv loader using "|" as a separator. This still left a lot of information to clean. Regular expressions were used to remove dashes and NAs rows.
```{r}
suppressWarnings(library(stringr))
suppressWarnings(library(dplyr))
library(ggplot2)
library(DT)

#Reading the Tournament file and separing by | and omiting the NAs
data = read.csv(file = "https://raw.githubusercontent.com/dapolloxp/data607/master/project1/tournamentinfo.txt", header = FALSE, sep = "|",na.strings = c(" ", ""),stringsAsFactors = FALSE)[1:10] %>% na.omit()

#Spliting the Players and merging rows
data.new <- cbind(data[seq(1, nrow(data), by = 2 ), ],
                data[seq(2, nrow(data), by = 2), 1:2 ])[-1,]
#Naming the column
colnames(data.new) <- c("num","Name","Points","round_1","round_2","round_3","round_4","round_5","round_6","round_7","State","Rating")

#Extracting the pre rating 
data.new$Rating <- as.numeric(as.character(str_remove_all(str_extract_all(data.new$Rating,"R: \\s?([\\d]{3,4})"),"R: \\s?")))
#Removing the letters W D L 
data.new <- data.frame(lapply(data.new, gsub, pattern = "([W D L])[ ]{2,3}(\\d+)", replacement = "\\2"), stringsAsFactors = FALSE)
data.new$Points <- as.numeric(data.new$Points )
data.new$Rating <- as.numeric(data.new$Rating)
#Calculating the opponent average 
data.new$Average <- apply(data.new[,4:10], MARGIN=1,function(x) {
   suppressWarnings(round(mean(as.numeric(as.character(data.new$Rating[as.numeric(x)])),na.rm =TRUE)))
    })
```
# DataTable
```{r DataTable}
#creating the DataTable with preferenced column
datatable(data.new[ c(2,3,11,12,13)]) 


#creating the CSV file
write.csv(data.new[ c(2,3,11,12,13)], file = "Data.csv")
```
#  Data Selection and Visualization

```{r Data Graph}
#looking at the graph its seem fairly normal distrubution.
ggplot(data=data.new) + geom_histogram (aes(x=Points),color="Blue", fill="white", bins = 6) + labs(title="Points Frequency plot",x="Points", y = "Frequency ") 


```