---
title: "Data607 Assignment"
author: "David Apolinar, Anthony Muñoz"
date: "2/14/2019"
output: html_document
---
```{r install packages}
library(stringr)
```


### Problem 3 Both

#### 3-1
```{r 3-1}

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
regex <- "([a-zA-Z,. ]){2,}"
name <- unlist(str_extract_all(raw.data, regex))
name
str(name)
data.name <- str_remove_all(name,"(\\w+)\\.\\s")
data.name
data.name <-gsub("(\\w+)\\,\\s(.*)","\\2 \\1", data.name)
df <- data.frame(str_split_fixed(data.name, " ", 2))
colnames(df) <- c("First Name","Last Name")
df

```
#### Problem 3-2
#### Checking whether the names have a title
```{r 3-2}
str_detect(name, "[A-Za-z]{2,3}\\.")
# We can see that 3rd and last name contain titles
unlist(str_extract(name, "[A-Za-z]{2,3}\\."))

```
#### Problem 3-3
#### Check whether a character has a 2nd name
```{r 3 c}
# We need to remove the title first
cleaned_names <- str_remove_all(name,"[A-Za-z]{2,3}\\.")
cleaned_names
# let's detect whether any of these have a more than 3 names
str_count(str_trim(cleaned_names), "\\w+") > 2
## Only C. Montomery Burns has a middle name


```
### Problem 4 Both
#### a. [0-9]+\\$  this will extract any digits, ending with a '$'
```{r 4 a}
#this will extract any digits, ending with a ampersand
unlist(str_extract_all(c("test23455$eklelkd8989k"), "[0-9]+\\$"))

```
#### b. \\b[a-z]{1,4}\\b

#### This will extract and lower case words with a break, less than 4 characters.
```{r 4 b}
unlist(str_extract_all(c("This will extract And lower Case words with a break, less than 4 characters."), "\\b[a-z]{1,4}\\b"))
```
#### c .*?\\.txt$
#### This will extract files that end with a .txt extension
```{r 4 c}
unlist(str_extract("this sentence contains a filename called t.txt with a .txt extension filename.txt", ".*?\\.txt$"))
paste("The full filename will be extracted because the regex pulls until the ending .txt")
unlist(str_extract("file with a .txt in the name.txt", ".*?\\.txt$"))
```
#### d. \\d{2}/\\d{2}/\\d{4}
#### This will extract a date pattern, such as 02/19/2018
```{r 4 d}
message <- "This contains 07/04/1776, such as 3494095, but it also contains, such as 33/35/3533444"
paste("The following text will be extracted: ")
unlist(str_extract_all(message, "\\d{2}/\\d{2}/\\d{4}"))
```


#### e. <(.+?)>.+?</\\1>
#### This captures anything that starts with a <tag> anything in between, up to the closing tag</tag>. Anything after is not captured.
```{r 4 e}
unlist(str_extract_all("Not captured <html><b><i>this is captured</i></html> This is not captured", "<(.+?)>.+?</\\1>"))
```

### Problem 9 David
####Find the hidden message
```{r problem 9 David}
hiddenMessage <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
### Showing method below for how I figured it out
# extract letters
unlist(str_extract_all(hiddenMessage, "[[:alpha:]]"))
# extract upper
unlist(str_extract_all(hiddenMessage, "[A-Z]"))
## looks like a sentence, let's extract at the periods
new.hidden <- unlist(str_extract_all(hiddenMessage, "([A-Z\\.])"))
# let's get rid of the sapces
new.hidden <- str_c(new.hidden, collapse = "")
## let's split at the periods
new.hidden
## replace periods
final.msg <- gsub("\\.", " ", new.hidden)
tolower(final.msg)


```
### Problem 9 Anthony
#### Find the Hidden Message
```{r Anthony}

data <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo\nUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO\nd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5\nfy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

data <- paste(unlist(str_extract_all(data,"[A-Z]")), collapse = "")
gsub("(.{15})(.{3})(.{3})(.{1})(.{5})(.{4})","\\1 \\2 \\3 \\4 \\5 \\6", data)


```

