---
title: "Assignment 9"
author: "David Apolinar"
output: 
  html_document:
   toc: true
   toc_float: true
---

  
### Intro {.tabset .tabset-fade .tabset-pills}

For this assignment, I decided to try out 3 of NYtimes APIs. I used a few different parameters for the Articles and Newswire APIs to show the different results that can be returned. 

* The Articles API
* The Movie Reviews API
* The Times Newswire API 

All of the data was loaded using the jsonlite package, converted as a dataframe, and flatten out to make table selection simplier.

```{r library loads}
library(httr)
library(ggplot2)
library(knitr)
library(jsonlite)
library(magrittr)
library(dplyr)
library(DT)

```   
 
### NYTimes Article API

Using this API, I decided to return all articles that contain keywords for the following topics:

* Microsoft
* Telsa
* Amazon

Searching all articles related to Microsoft and returning select columns:

Note: Column names were renamed to make them user friendly.

```{r}
api_file <- "/Users/davidapolinar/Dropbox/APIkeys/nytimes.txt"
key<-read.delim(api_file, header = FALSE)
keyword <- "microsoft"

query <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?", "q=",keyword,"&api-key=",key$V1 ,sep = "")
y <- fromJSON(query, flatten = TRUE) %>% data.frame()

out <- y %>% select(response.docs.web_url, response.docs.section_name, response.docs.headline.main, response.docs.byline.original) %>% rename(URL = response.docs.web_url, SectionName = response.docs.section_name, Headline = response.docs.headline.main, Author=response.docs.byline.original)
DT::datatable(out, options = list(pageLength = 5))
```

Searching all articles related to Telsa and returning select columns:

```{r}
api_file <- "/Users/davidapolinar/Dropbox/APIkeys/nytimes.txt"
key<-read.delim(api_file, header = FALSE)
keyword <- "tesla"

query <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?", "q=",keyword,"&api-key=",key$V1 ,sep = "")
y <- fromJSON(query, flatten = TRUE) %>% data.frame()

out <- y %>% select(response.docs.web_url, response.docs.section_name, response.docs.headline.main, response.docs.byline.original) %>% rename(URL = response.docs.web_url, SectionName = response.docs.section_name, Headline = response.docs.headline.main, Author=response.docs.byline.original)
DT::datatable(out, options = list(pageLength = 5))
```

Searching all articles related to Amazon and returning select columns:

```{r}
api_file <- "/Users/davidapolinar/Dropbox/APIkeys/nytimes.txt"
key<-read.delim(api_file, header = FALSE)
keyword <- "amazon"

query <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?", "q=",keyword,"&api-key=",key$V1 ,sep = "")
y <- fromJSON(query, flatten = TRUE) %>% data.frame()

out <- y %>% select(response.docs.web_url, response.docs.section_name, response.docs.headline.main, response.docs.byline.original) %>% rename(URL = response.docs.web_url, SectionName = response.docs.section_name, Headline = response.docs.headline.main, Author=response.docs.byline.original)
DT::datatable(out, options = list(pageLength = 5))
```


### NYTimes Movie Review APIs

This API returns all movie reviews by the NYtimes author. The APIs can be customized to return several variables depending on what is selected. For demo purposes, I kept it to just the title, headline, and author.

```{r}
query2 <-paste0("https://api.nytimes.com/svc/movies/v2/reviews/search.json?", "api-key=",key$V1 ,sep = "")
x <- fromJSON(query2, flatten = TRUE) %>% data.frame()
results1 <- x %>% select(Title = results.link.suggested_link_text, Headline = results.headline, Author = results.byline)
datatable(results1, options = list(pageLength = 5))
```

### NYTimes Newswire API

This API gets all of the latest articles published. For this section, I used a keyword for the relevant sections that I was interested in retrieving.

Getting all the latest articles related to Business

```{r}
source = "all"
section = "Business"

query3 <-paste0("https://api.nytimes.com/svc/news/v3/content/", source, "/", section, "?api-key=",key$V1 ,sep = "")
x <- fromJSON(query3, flatten = TRUE) %>% data.frame() 
results3<-x %>% select(results.byline, results.section, results.title) %>% rename(Author=results.byline, Section=results.section, Title=results.title)
datatable(results3, options = list(pageLength = 5))
```

Getting all the latest Opinion articles

```{r}
source = "all"
section = "Opinion"

query3 <-paste0("https://api.nytimes.com/svc/news/v3/content/", source, "/", section, "?api-key=",key$V1 ,sep = "")
x <- fromJSON(query3, flatten = TRUE) %>% data.frame() 
results3<-x %>% select(results.byline, results.section, results.title) %>% rename(Author=results.byline, Section=results.section, Title=results.title)
datatable(results3, options = list(pageLength = 5))
```